### 2023-09-19 presentation/siesta-2023.qmd
feat(presentation): init siesta 2023 presentation

---
title: "Bridging the Gap between Visual and Analytical ML Testing"
subtitle: "SIESTA 2023"
date: 2023-09-11
mainfont: "Iosevka Aile"
monofont: "Iosevka"
affiliations:
  - id: tud
    name: "Delft University of Technology"
author:
  - name: "[***Aru***]{.underline}moy Shome"
    affiliation:
      - ref: tud
  - name: "Luis Cruz"
    affiliation:
      - ref: tud
  - name: "Arie van Deursen"
    affiliation:
      - ref: tud
format:
  revealjs:
    theme: [default, custom.scss]
license: "CC BY"
copyright:
  holder: "Arumoy Shome"
  year: 2023
bibliography: siesta-2023.bib
---

## 🤫🤫🤫🤫

::: {.callout-tip title="Towards Understanding Machine Learning Testing in Practice [@shome2023towards]"}
:::

::: {.callout-important title="Discretion is advised!"}
~~Some~~Most parts of this presentation are unpublished!
:::

## ML Testing

::: {.callout-tip title="Machine Learning Testing: Survey, Landscapes, and Horizons[@zhang2020machine]"}
:::

::: {.fragment}
![ML lifecycle](ml-lifecycle.svg){#fig-ml-lifecycle}
:::
## From Implicit Expectations...

:::: {.columns}

::: {.column width="60%"}
![](histogram.jpg){width="50%"}

![](qqplot.png){width="50%"}
:::

::: {.column width="40%" .fragment}
+ Tend to be messy/quick/hacks
+ Exploratory, data understanding/sense making
+ Domain expertise and knowledge of practitioner *embedded* in visualisations
:::
::::
## ...To Explicit Tests
:::: {.columns}
::: {.column width="40%"}
+ Verifiable across entire ML lifecycle
+ Scales across org/team change
+ Scales to large-scale production systems and big data
+ Eg. *Shapiro-Wilk Test* of normality
:::

::: {.column width="60%" .fragment}
```python
import numpy as np
from scipy import stats

x = stats.norm.rvs(size=100)
shapiro_test = stats.shapiro(x)

assert shapiro_test.pvalue <= 0.05
```
:::
::::

## The Hunt for Data Properties {.smaller}

:::: {.columns}
::: {.column width="50%"}
**Literature**

+ Review of conferences and journals focusing on visualisations: HCI,
SE, VIZ, ML & DMM
+ Review of grey literature
:::

::: {.column width="50%"}
**Computational NBs**

+ Kaggle
+ Github
+ @pimentel2019large & @quaranta2021kgtorrent
:::
::::

## Here Come the Numbers {.smaller}

::: {.callout-important title="1.7K"}
Notebooks from Github with an `assert` statement
:::

::: {.callout-warning title="337"}
Notebooks with related `assert` & visualisation pairs
:::

::: {.callout-note title="90"}
Notebooks analysed in full depth
:::

::: {.callout-tip title="22"}
Examples of `assert` & visualisation pairs
:::

# Results

## Area under ROC curve {.smaller}

:::: {.columns}
::: {.column}
![](select-09.png)
:::

::: {.column}
```python
assert roc_auc_score(
    y_test, wv_model.predict_proba(X_test_wv)[:, 1]
) > 0.92,
"something's wrong with your features"
```
:::
::::
## Loss, convergence {.smaller}

:::: {.columns}
::: {.column}
![Type 1](select-12.png){#fig-left}
:::

::: {.column}
![Type 2](select-267.png){#fig-right}
:::
::::

:::: {.columns}
:::{.column}
```python
assert np.mean(history[:10]) >np.mean(history[-10:]),
"RNN didn't converge"
```
:::

::: {.column}
```python
# assert we actually descended at each step
for i in range(1, n_steps):
    assert (
        loss_array_true_grad[i] - loss_array_true_grad[i-1]
    ) <= 0
```
:::
::::

::: {.callout-warning title="Smell"}
The assertion for @fig-right is brittle compared to @fig-left.
:::
## Decision boundary & accuracy of classifier {.smaller}

:::: {.columns}
::: {.column}
![](select-04.png)
:::

::: {.column}
```python
assert accuracy_score(y_test, pred) >0.95
```
:::
::::

::: {.callout-note title="Open question"}
Some visualisations are more complex compared to their counter-part
assertion. Is it because we don't yet know how to write better
assertions for such visualisations?
:::
## Hyper-parameter grid search

:::: {.columns}
::: {.column}
![](select-290.png){width="60%"}
:::

::: {.column}
::: {.callout-note title="Qualitative analysis of hyper parameters"}
The author is visually checking all combinations of two hyper
parameters `C` and `gamma` to identify the combinations that cause the
model to {under,over}fit.
:::
:::
::::

# Our objective is to [*recommend*]{.underline} analytical tests automatically {.r-text-fit}

## Thank You

::: {layout-nrow="2"}
![](qr-cat.jpeg)

![](qr-lugano.jpeg)

![](qr-illustrate.jpeg)

![](qr-picaso.jpeg)
:::
## Refs {.smaller}