### 2023-09-19 catalogue/catalogue.qmd
feat(catalogue): init catalogue

---
title: Visualisation, assertion catalogue
author: Arumoy Shome
mainfont: "Iosevka Aile"
monofont: "Iosevka"
monofontoptions:
  - Scale=0.9
format: pdf
---

## (04) SVM decision boundary, accuracy
![](select-04.png)

```python
assert accuracy_score(y_test, pred) > 0.95
```

## (05) input image, accuracy
![](select-05.png)

```python
assert accuracy > 0.85, "Poor accuracy {:.3f}".format(accuracy)
```

## (09) roc, auc
![](select-09.png)

```python
assert roc_auc_score(y_test, wv_model.predict_proba(X_test_wv)[:, 1]) > 0.92,
"something's wrong with your features"
```

## (12) loss, convergence
![](select-12.png)

```python
assert np.mean(history[:10]) >np.mean(history[-10:]), "RNN didn't converge"
```

::: {.callout-note}
This is the most common visualisation,assertion pair.
:::

## (13) lineplot, correlation
![](select-13.png)

```python
assert answer_nine() >= -1. and answer_nine() <= 1.,
"Q9: A valid correlation should between -1 to 1!"
```

# (35) input image, `predict_proba`

![](select-35a.png)

```python
np.testing.assert_almost_equal(pred_proba, 2.0355723e-05, decimal=3)
```

![](select-35b.png)

```python
np.testing.assert_almost_equal(pred_proba, 0.9988895, decimal=3)
```

::: {.callout-warning title="Smell"}
How did the author identify the probability threshold? Perhaps
comparision with ge/le is better?
:::

# (47) log-scale lineplot, coefficient & intercept

![](select-47.png)

```python
assert l.coef_ < 0, "Curve must be decreasing"
assert l.intercept_ > 1, "Curve must start higher"
assert l.score(np.log(X).reshape((-1,1)), np.log(values)) > 0.9, "Unusually large error."
```

::: {.callout-caution}
We need to revisit this one!
:::

# (61) various image rescaling techniques

::: {layout-ncol=3}
![Original input image](select-61a.png)

![Rescaled image](select-61b.png)

![Decomposed image](select-61c.png)
:::

# (63) visual and analytical assertion on mean recall

![](select-63.png)

```python
assert np.mean(dev_recall_history[-10:]) > 0.85, "Please train for at least 85% recall on test set. "\
                                                  "You may need to change vectorizer model for that."
```

# (65) Scaling and Normalization

![](select-65a.png)

```python
assert scaled_data[scaled_data.argmax()] <= 1
assert scaled_data[scaled_data.argmin()] >= 0
```

::: {.callout-warning title="Smell"}
Perhaps the visualisation is redundant?
:::

![](select-65b.png)

```python
assert normalized_data[normalized_data.argmax()] < 10
assert normalized_data[normalized_data.argmin()] > -7
```

::: {.callout-warning title="Smell"}
How did the author identify the thresholds?
:::

::: {.callout-tip title="Visual testing"}
This is the most trivial example of visual testing - checking the
distribution of features.
:::

# (78) confusion matrix, false positive/negative

![](select-78.pdf){width="50%"}

```python
assert conf_mat.shape == (2,2)
assert conf_mat[0][1] < 7
assert conf_mat[1][0] < 3
```

::: {.callout-warning title="Smell"}
The values should be divided by $n$ (total number of observations) to
account for bias in the dataset.
:::

# (82) DNN mean reward history

![](select-82.png)

```python
assert np.mean(mean_rw_history[-10:]) > 10.
```

# (86) lolopy performance testing

![](select-86.png)

```python
lolopy_timing = interp1d(results['n'], results['lolopy_train'])
lolo_timing = interp1d(results['n'], results['scala_train'])
slowdown = lolopy_timing(100) / lolo_timing(100)
print('Training slowdown: {:.2f}'.format(slowdown))
assert slowdown < 2
```

::: {.callout-caution}
This is more devops/mlops imo, do we include in this our current
paper?
:::

# (88) histograms for outlier detection

![](select-88a.png)

::: {.callout-tip title="Visual testing"}
A very nice example of a visualisation to determine/understand the
property of the underlying dataset. Check the NB for more details.
:::

![](select-88b.png)

::: {.callout-warning title="Smell"}
The author identifies outliers in the dataset using histograms.
Typically a boxplot is better for outlier detection?
:::

# (110) data property from domain expertise

![](select-110.png)

```python
# In the plot it looks like there are no exceptions, but let's query
the dataset to be 100% sure
days_with_more_than_6_entries_per_day = df_daily_data.groupby('datum').filter(lambda x: len(x) > 6).index

assert len(days_with_more_than_6_entries_per_day) == 0
```

# (117) time-series spot checking

::: {layout-ncol=2}
![Spot checking $y$, $x$](select-117a.png){#fig-117a}

![Check error of prediction](select-117b.png){#fig-117b}
:::

In @fig-117a the author spot checks the values of $y$ at specific values
of $x$.

```python
# y_t[i] are the number of cases on the i-th day
assert y_t[0] == 85
assert y_t[10] == 1452
assert y_t[800] == 65032
```

::: {.callout-warning title="Smell"}
It is difficult to identify the exact value of $y$ from a
visualisation. When the data changes, the assert will fail right away.
:::

In @fig-117b the author evalues the performance of a predictor using a
custom measure/metric. Performance also visually checked using a
lineplot.

```python
# Result of the wfc function for the best choice of s
assert np.isclose(best_wfv, 7685.546728971963, 0.001)
```

# (143) input image, `predict_proba`

::: {layout-nrow=2}
![](select-143a.png)

![](select-143b.png)

![](select-143c.png)

![](select-143d.png)
:::

```python
assert 0 <= predict_image(face_model, TP_URL).first().prediction <= 1
assert 0 <= predict_image(face_model, TN_URL).first().prediction <= 1 
assert 0 <= predict_image(face_model, FP_URL).first().prediction <= 1
assert 0 <= predict_image(face_model, FN_URL).first().prediction <= 1
```

::: {.callout-warning title="Smell"}
The assertions are on the theoretical limits of probability.
:::

![](select-143g.png)

::: {.callout-tip title="Visual Testing"}
The author is visualising the decision boundary of a MLP along with the
accuracy of the model. 
:::

# (152) visual & statistical distribution check

![](select-152a.png)

```python
for feature in range(data_transformed.shape[1]):
    assert kstest(data_transformed[:, feature], 'norm').statistic < 1e-2
```

![](select-152b.png)

::: {.callout-tip title="Visual Testing"}
We need to revisit this, but I think the author is trying to validate if the
output generated by the GAN is similar to the original data.
:::

# (153) bootstrap estimate of mean & variance

::: {layout-ncol=2}
![Mean](select-153a.png){#fig-153a}

![Variance](select-153b.png){#fig-153b}
:::

The assertions for Figure @fig-153a and @fig-153b are the following.

```python
assert 80 < boot_mean_mean < 81
assert 0.3 < boot_var_mean < 0.5
```

```python
assert 7 < boot_mean_std < 10
assert 0.23 < boot_var_std < 0.29
```

::: {.callout-warning title="Smell"}
While the mean of the sample can be identified from the histogram, I don't
think the same holds for the variance?
:::

# (165) pie chat

![](select-165.png)

```python
assert_equal(num_passengers_by_sex.index.tolist(), ["female", "male"])
```

# (172) image augmentation

::: {layout-ncol=2}
![Original image](select-172a.png)

![augmented image](select-172b.png)
:::

```python
assert not np.allclose(x_aug, x_original)
```

# (202) input image and predictions

![](select-202a.png)

::: {.callout-tip title="Visual Testing"}
This is a fairly common example of visual testing where the image input along
with the prediction of the model is visualised.

Although we can use various performance metrics (accuracy, precision, recall,
etc), this form of qualitative testing is still prominent when working with
images.
:::

# (204) logistic regression

![](select-204a.png)

::: {.callout-tip title="Visual Testing"}
I think the author is trying to visualise the results of the optimization here? Need second opinion.
:::

![](select-204b.png)

```python
ans4 = compute_loss(X_expanded, y, w)
np.testing.assert_almost_equal(ans4, 0.3042764)
```

::: {.callout-warning title="Smell"}
Better to use ge/le comparison rather than hardcoded threshold?
:::

# (210) great expectations

![](select-210.pdf)

```python
result = solar.expect_column_values_to_be_between(field_map["date"], 
                                                    solar[field_map["date"]].min(), 
                                                    pd.Timestamp.now(),
                                                    parse_strings_as_datetimes=True)
```

# (228) linear approximation using domain knowledge

::: {layout-ncol=2}
![](select-228a.png){#fig-228a}

![](select-228b.png){#fig-228b}
:::

For Figure @fig-228a.

```python
assert np.allclose(y_approx.mean(), 0)
assert np.allclose(y_approx.std(), 0.6721265191518054)
```

For Figure @fig-228b.

```python
assert np.allclose(y_long.mean(), 0.2742076173424)
```

::: {.callout-warning title="Smell"}
Its not possible to identify the mean of a distribution from the visualisation to such precision. I think a ge/le comparison is better.
:::

# (267) loss is decreasing

![](select-267.png)

```python
# assert we actually descended at each step
for i in range(1, n_steps): 
    assert (loss_array_true_grad[i] - loss_array_true_grad[i-1]) <= 0
```

::: {.callout-warning title="Smell"}
The author checks that the loss is decreasing at each step/epoch. How do we ensure that we are not stuck in a local minimum with this assertion style? Perhaps its better to keep it general and only check if the last $n$ observations is lower than the first $n$ observations?
:::

# (275) loss, accuracy

![Loss](select-275.png)

```python
acc = compute_test_accuracy(model)
assert acc>0.96, "Bidirectional RNNs are better than this!"
```

::: {.callout-caution title="Luis"}
The visualisation is of the loss, but the assertion is on the accuracy. They feel "kind of" related? What to do here?
:::

# (290) hyper-parameters

![](select-290.png)

::: {.callout-tip title="Visual testing"}
Qualitative assessment of hyper-paramer configuration that causes
over-fitting in model.
:::

# (332) recurrent neural networks

![True vs. predicted labels](select-332a.png){#fig-332a}

```python
r2_gru = r2_score(y_test, y_pred)
assert r2_gru > 0.6
```

::: {.callout-note}
The $R^2$ or coefficient of determination is derived from the predicted and actual labels.
:::