### 2023-10-12 report/icst24.tex
feat(report): 2023-10-12 init implications section

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{array}
\usepackage{framed}
\newcommand{\highlight}[1]{\begin{framed}%
  \noindent\emph{#1}
\end{framed}}

\definecolor{backcolor}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
  basicstyle=\ttfamily\footnotesize,
  keywordstyle=\color{blue},
  stringstyle=\color{red},
  commentstyle=\color{green},
  breakatwhitespace=true,
  breaklines=true,
  keepspaces=true,
  showspaces=false,
  showstringspaces=false,
  showtabs=false,
  tabsize=4,
  captionpos=b,
  numbers=left,
  numberstyle=\tiny\color{gray},
  backgroundcolor=\color{backcolor},
}
\lstset{style=mystyle}
\usepackage{graphicx}
\usepackage{caption}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\begin{document}

\title{Bridging the Gap Between Visual and Analytical Machine Learning
  Testing}

\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
}

\maketitle

\begin{abstract}

Testing ML systems is a highly interactive process which demands a human-in-the-loop approach. In addition to writing tests for the code base, practitioners are required to analyse and interpret several visualisations using their domain expertise to validate if an ML system satisfies the required set of functional and non-functional properties. Visualisations are frequently used to qualitatively assess various parts of an ML pipeline. However, implicit knowledge gained from visualisations must be translated to explicit analytical tests that fail when there is change in any component of the ML pipeline. We conduct an empirical analysis of Jupyter notebooks to catalogue the state-of-the-art mappings between ML visualisations and assertions. We mine Github to collect Jupyter Notebooks that contain assertions written in Python. We develop a novel methodology to identify 1764 notebooks which contain an assertion below a visualisation. We manually analyse the 1.7K notebooks and identify 551 notebooks where at least one pair of visualisation and assertion are semantically related to one another. We further read and understand 90 notebooks in depth to identify 26 qualitative and their corresponding quantitative testing patterns. We showcase five such patterns in this paper and plan to make the entire catalogue publicly available in the near future.

% TODO what were the primary findings of this study?

\end{abstract}

\begin{IEEEkeywords}
  ML Testing, Visualisations, ???
\end{IEEEkeywords}

\section{Introduction}\label{sec:intro}

Recent years show a shift from chasing high correctness of ML models to ensuring that they are safe for application in safety-critical domains such as healthcare, finance and autonomous driving. With the advent of Large Language Models, the power of AI to directly influence the daily lives of humans is growing at a radical pace. As reports of ``hallucinated'' facts, racist predicts and autonomous car crashes steadily increase, ensuring that ML systems function as they were intended to, takes precedence.

Visualisations are used throughout the ML development lifecycle to test various properties of the ML system. In the early stages of the ML development lifecycle, visualisations are extensively used to make sense of the data and verify its statistical properties. During the model development phase, visualisations are used to summarise metrics and contrast different learning algorithms and iteratively fine-tune ML models. Once the model is deployed in production, visualisations are used to continually monitor their performance and trigger a new training cycle once their performance drops below a certain threshold.

Building ML systems is highly iterative and experimental. Information gained from ML visualisations is used to make design and implementation decisions for the following steps of the ML pipeline. For instance, we may visualise the distribution of our training data and find that it is normally distributed. Based on this information, we may opt for a Linear Regression model which assumes normality in the underlying data. However, such expectations regarding the data may be violated once the ML system is deployed in production, where the data constantly changes as a reflect of the real world. Implicit expectations obtained from visualisations must therefore be translated to analytical tests that fail once our expectations regarding the ML system are no longer satisfied.

In contrast to prior work done in ML testing, we approach this pressing topic from a novel direction. We conduct an empirical analysis of computational notebooks obtained from Github, to understand the process of testing ML systems in practice. In particular, we focus on the combination of a qualitative form of testing (using visualisations) and a quantitative form of testing (using analytical assertions). The research questions along with the contributions of this paper are as follows.

\begin{description}
\item[RQ1.] How frequently do ML practitioners formulate analytical tests from the visualisations they create to test ML systems?

We mine 54K Jupyter notebooks from Github that contain an assertion written in Python. We develop an automated technique to identify 1764 notebooks with an assertion below a visualisation. We manually analyse the 1.7K notebooks and identify 551 notebooks that contain at least one pair of visualisation and assertion that are semantically related to each other.

\item[RQ2.] What are the types of visualisation and assertion pairs typically used by ML practitioners when testing ML systems?

From the 551 notebooks, we further perform an in-depth analysis of 90 notebooks. We identify 26 unique visualisation and assertion pairs. We discuss five such pairs in this paper and plan to release the full catalogue publicly in the near future.

\end{description}

% TODO what are the primary findings of this paper...

The paper is organised as follows. Section~\ref{sec:prelim} summarises the related literature and technical information relevant to this study. The methodology for the empirical study is explained in Section~\ref{sec:method} and the results of the study are presented in Section~\ref{sec:result}. The paper concludes with a discussion on the implications of our results (Section~\ref{sec:discuss}), threats to validity (Section~\ref{sec:threats}) and future work (Section~\ref{sec:conclude}).

\section{Preliminaries}\label{sec:prelim}

% TODO how is our work different from prior work?

This section summarises the body of prior scientific work that has been done in ML Testing, computational notebooks and visual analytics. The section concludes with an overview of technical knowledge required for this paper.

\subsection{ML Testing}\label{sec:ml-testing}

Testing ML enabled software systems poses more challenges compared to their traditional counter part. While traditional software systems mature through change in their codebase, ML enabled systems experience change in the training data and the machine learning model in addition to the codebase\cite{CITEME}. With ML being adopted into safety-critical domains that can affect human lives, ensuring that ML systems are correct, robust to data perturbations and not biased by race, colour or gender is of paramount importance. Existing scientific literature on ML testing broadly focuses on two aspects. First, on functional properties such as correctness and robustness of the model towards unseen data. And second, on non-functional properties such as fairness, interpretability and privacy.

To test the correctness of ML models, several improvements over existing test adequacy metrics have been proposed. Tools such as \textit{DeepExplore} and \textit{DeepGauge} propose new test adequacy metrics such as neuron coverage adapted for ML enabled software systems~\cite{pei2017deepexplore, ma2018deepgauge, gerasimou2020importance}. Formal verification methods have also been proposed that try to provide formal guarantee of robustness against adversarial examples. Such methods however are only feasible for statistical ML models and become computationally expensive for more complex models such as deep neural networks~\cite{zhu2021deepmemory, baluta2021scalable}. Several works have been conducted on generating and detecting adversarial inputs for ML models. Data augmentation techniques based on fuzzing, search based software testing and mutation testing have been proposed to generate adversarial examples that can be used during model training to improve its robustness~\cite{braiek2019deepevolution, gao2020fuzz, wang2021robot, zhang2020white}. It is however not possible to include all variations of adversarial examples into the training data. Thus, methods have been proposed to detect adversarial inputs during runtime~\cite{xiao2021self, wang2020dissector, wang2019adversarial, berend2020cats}.

% TODO summarise literature on fairness testing
%
Bulk of the contributions on testing non-functional properties of ML systems, focuses on fairness.

\subsection{Computational Notebooks and Software
  Engineering}\label{sec:notebooks}

Computational notebooks have been ubiquitously adopted by the machine learning community for developing ML enabled systems. Although originally intended to promote reproducible software, computational notebooks are far from being reproducible due to lack of software engineering best practices such as separation of concern, testing and versioning~\cite{pimentel2019large}.

\cite{psallidas2019data} provide an overview of the evolving landscape of computational notebooks by analysing six million Python notebooks, two million enterprise data science pipelines, and source code and metadata from over 900 releases of 12 important data science libraries. Their findings can be used by system builders for improving data science tools and also by data science practitioners to understand the current trends and what technologies they should focus on. \cite{pimentel2019large} mined 1.4 million notebooks from Github to conduct an empirical study on the good and bad coding practices in computational notebooks. Based on their analysis, the authors propose guidelines on improving reproducibility of computational notebooks. To enable future research on computational notebooks, \cite{quaranta2021kgtorrent} mine Kaggle~\footnote{https://kaggle.com} and present \textit{KGTorrent}, a public dataset consisting of approximately 2.5 million Jupyter notebooks written in Python. The dataset also contains a relational database dump of metadata regarding publicly available notebooks on Kaggle.

Studies with human subjects have been conducted to gain a deeper understanding of the challenges faced by ML practitioners when developing ML models inside notebooks. The results indicate that ML practitioners work in a highly iterative fashion, often experimenting with multiple strategies to analyse the data or produce meaningful visualisations~\cite{kandel2012enterprise, kery2018story, liu2019understanding, chattopadhyay2020wrong}. Studies have also been conducted to understand how practitioners generate, evaluate and manage alternative hypothesis, visual designs, methods, tools, algorithms and data sources to arrive at the final implementation~\cite{liu2019understanding,kandel2012enterprise}. The findings from these studies can be used as guidelines for improving existing notebook technologies or designing new ones.

To manage and prune the multiple versions of code that accumulate over time when developing in notebooks, \cite{head2019managing} propose a code gathering tool that allows practitioners to review and only keep the relevant version of code. Other tools such as \textit{WrangleDoc} and \textit{VizSmith} have also been proposed to aid ML practitioners when working within computational notebooks~\cite{yang2021subtle, bavishi2021vizsmith}.

\subsection{Visualisations and Machine Learning}\label{sec:visualisations}

As ML augment software systems in safety-critical domains, emphasis has been put into explainability of ML models. ML models and the underlying data is complex and multi-dimensional. To combat the ``curse of dimensionality'', visual analytics has been widely adopted by the ML community to understand the data and the internal workings of ML models.

Prior studies have been conducted to understand how visual analytics techniques are currently being used in ML. \cite{yuan2021survey} conduct a systematic review of 259 papers and propose a taxonomy of visual analytics techniques for ML. \cite{hohman2019visual} conduct a survey of visual analytics techniques for Deep Learning Models. The findings of the study indicate that visual analytics has been widely adopted in ML for model explanation, interpretation, debugging and improvement.

Several tools have been proposed by the visual analytics research community to aid practitioners in understanding how their ML models operate. \cite{wexler2020if} propose \textit{What-If}, a visual analytics tool to explore alternative hypotheses, generate counterfactual reasoning, investigate decision boundaries of the model and how change in data affects the model predictions. To reduce the cognitive load of ML practitioners during model building phase, \cite{amershi2015modeltracker} propose \textit{ModelTracker}. Given a trained model and a sample dataset, ModelTracker presents all the information in traditional summary statistics along with the performance of the model.

\textit{ESCAPE}, \textit{GAM Coach}, \textit{Angler} and \textit{Drava} are a few other tools that have been proposed to handle specialised use-cases such as identifying systemmatic errors in ML models, generating counterfactual explanations for Generalised Additive Models, prioritising machine translation model improvements and relating human concepts with semantic dimensions extracted by ML models during disentangled representation learning\cite{ahn2023escape, wang2023gam, robertson2023angler, wang2023drava}.

\subsection{Internal Structure of Jupyter Notebooks}\label{sec:nbformat}

\begin{lstlisting}[caption={Top-level JSON format of Jupyter notebooks. See Listing~\ref{lst:cell-md},~\ref{lst:cell-code} and~\ref{lst:cell-output} for internal structure of various cells.}, label={lst:nbformat}]
{
  "metadata": {
    "kernel_info": {...},
    "language_info": {...},
  },
  "nbformat": 4,
  "nbformat_minor": 0,
  "cells": [ # dictionary of all cells in notebook ]
}
\end{lstlisting}
\begin{lstlisting}[caption={JSON structure of markdown cells.}, label={lst:cell-md}]
{
  "cell_type": "markdown",
  "metadata": {},
  "source": [
    "# Hello world example\n",
    "The following cell contains python code for greeting people!\n"
  ]
}
\end{lstlisting}

\begin{lstlisting}[caption={JSON structure of code cells.}, label={lst:cell-code}]
{
  "cell_type": "code",
  "metadata": {},
  "source": [
    "def greet(name):\n",
    "    print(f'Hello, {name}!')\n"
  ]
}
\end{lstlisting}

\begin{lstlisting}[caption={JSON structure of code cells with an image output.}, label={lst:cell-output}]
{
  "cell_type": "code",
  "metadata": {},
  "source": [...],
  "outputs": [
    {
      "output_type": "display_data",
      "data": {
        "image/png": "<base64-encoded-png-data>"
      }
    }
  ]
}
\end{lstlisting}

The data comprised within Jupyter Notebooks is stored in the JSON format, and thus can be parsed programmatically. Listing~\ref{lst:nbformat} shows the top-level json structure of Jupyter Notebooks, adopted from the official nbformat website.~\footnote{https://nbformat.readthedocs.io/en/latest/index.html}. For simplicity, we only show the parts that are relevant for this study and hide the irrelevant parts using ellipsis.

Notebooks contain four top level keys namely \texttt{metadata}, \texttt{nbformat}, \texttt{nbformat\_minor} and \texttt{cells}. The \texttt{matadata} key consists of additional information regarding the notebook such as the kernel and programming language. \texttt{nbformat} and \texttt{nbformat\_minor} contain the version of nbformat used in the notebook. The \texttt{cells} key consists of a list of all cells within the notebook.

Listings~\ref{lst:cell-md},~\ref{lst:cell-code} and ~\ref{lst:cell-output} present three examples of cells. Each cell is represented as a dictionary with the \texttt{cell\_type}, \texttt{metadata} and \texttt{source} keys. Cells can be of three types: ``markdown'', ``code'' or ``raw''. The contents of each cell can be found within the \texttt{source} key, as a list of strings.
Outputs generated by a code cell are stored as a list within the
\texttt{outputs} key as seen in Listing~\ref{lst:cell-output}. Static
visualisations are stored as base64 encoded png images under the
\texttt{"image/png"} key, which is nested within the \texttt{data} key
inside \texttt{outputs}.

\subsection{Post-Condition Checks}\label{sec:post-cond}

\begin{table*}
\centering
\caption{Frequent examples of post-condition checks observed in this study.}
  \begin{tabular}{m{0.3\textwidth} m{0.4\textwidth} m{0.3\textwidth}}
\hline
\emph{\textbf{Visualisation}}&
\emph{\textbf{Post-condition check}}&
\emph{\textbf{Rationale}}\\
\hline
\includegraphics[width=\linewidth]{post-cond-01.png}&
\begin{lstlisting}[language=Python]
assert f1.gca().has_data()
\end{lstlisting}&
The assertion checks if the output of the visualisation code cell contains an image.\\
\includegraphics[width=\linewidth]{post-cond-02.png}&
\begin{lstlisting}[language=Python]
assert_almost_equal(sigmoid(-50), 0, delta=1e-10)
assert_almost_equal(sigmoid(0), 0.5, delta=1e-10)
assert_almost_equal(sigmoid(50), 1, delta=1e-10)
\end{lstlisting}&
The assertions are spot-checking the sigmoid activation function.\\
\includegraphics[width=\linewidth]{post-cond-03.png}&
\begin{lstlisting}[language=Python]
# The axis should contain 5 bars
assert len(ax.containers[0]) == 5

# One of the x tick labels should be "blue"
tick_text = [tick.get_text() for tick in ax.get_xticklabels()]
assert "blue" in tick_text
\end{lstlisting}&
The asserts are validating the correctness of the visualisation itself.\\
\hline
\end{tabular}
\label{tab:post-cond}
\end{table*}

We found several instances of visualisation and assertion pairs which were realated to one another. However the assertions did not test a property of the ML system that was derived from the visualisation. Instead, the assertions were used to test the correctness of the code that produced the visualisation. Such assertions (and the corresponding visualisations) were excluded from this study. Table~\ref{tab:post-cond}provides examples of
post-condition checks which were frequently observed in this study.

\section{Methodology}\label{sec:method}

Figure~\ref{fig:method} presents an overview of all relevant steps used in this study. The empirical analysis comprises of three distinct phases namely: \textit{the data collection phase}, \textit{identification of related visualisation and assertion pairs} and finally a \textit{complete analysis of the notebooks} to identify the final set of 26 visualisation and assertion pairs.

The yellow boxes in the data collection phase represent filters that were applied to the initial corpus of 54K notebooks, to obtain the final sample size of 1.7K notebooks used in the empirical study. The red boxes represent the exclusion criteria used during phase two to remove notebooks note relevant for this study. Finally, the green boxes in phase 3 represent the manual analysis conducted to identify the final 26 visualisation and assertion pairs.

Boxes appearning next to one-another imply that they were applied using an "OR" operator. While boxes appearning below one-another imply an "AND" operator. Text in double quotes represent the string pattern used in the search query. Each phase of the methodology is explained in more detail below.

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{method.pdf}
  \caption{Methodology used to collect and analyse Jupyter notebooks
    in this study.}
  \label{fig:method}
\end{figure*}

\subsection{Data Collection}\label{sec:data-collect}

We mined public repositories from Github to collect $54,070$ Jupyter notebooks that contain the keyword ``assert'' in them. We use the Github advanced search syntax~\footnote{REFME} to isolate Jupyter Notebook. This translates to the following Github search query: \texttt{filename:"Jupyter Notebooks"}. Additionally, we include the ``assert'' keyword in the search query with an \texttt{AND} operator. The final search query is as follows: \texttt{filename:"Jupyter Notebooks" AND "assert"}.

To further reduce the sample size, we exclude notebooks that do not use popular ML libraries. We perform a string search to identify notebooks that import at least one of the ML libraries. The string patterns are derived from the python module name used in the import statements. The regex search query is as follows: \texttt{tensorflow|dask|torch|sklearn|scipy|numpy|pandas}.

We programmatically parse the JSON structure of the notebooks (more details regarding the internal structure of Jupyter notebooks is presented in Section~\ref{sec:nbformat}) and represent each notebook as a pandas dataframe for further analysis. We map each cell as a row in the dataframe. For each cell, we store the \texttt{cell\_type}, the \texttt{source} of the cell and the base64 encoded PNG string, into the corresponding columns of the dataframe. We exclude notebooks that do not contain any visualisations by checking if the image column of the dataframe is empty.

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{nb-structure.pdf}
  \caption{Arrangement of code cells containing assertion and
    visualisation.}
  \label{fig:cell-arrangement}
\end{figure*}

To identify visualisation and assertion pairs that are semantically related to one another, we only consider notebooks in which the visualisation and assertion code cells appear in a specific order. Figure~\ref{fig:cell-arrangement} presents a visual representation of various arrangements of code cells that may occur in notebooks. Note that the visual representation omits markdown cells since we only consider the order of code cells in our analysis. In the actual notebook however, one or more markdown cells may be present between the visualisation and assertion cells.

We conduct the manual analysis presented in Section~\ref{sec:identify-related-pairs} for a small random sample of notebooks containing each cell arrangement. We observed the signal-to-noise ratio for each cell arrangement and picked the arrangement with the highest signal-to-noise ratio. Table~\ref{tab:cell-arrangement} presents the cell arrangement in ascending order of signal-to-noise ratio with a summary of our observations.

Assertions defined in a cell above the visualisation cell, were typically not related to each other. Often, a markdown cell present between the assertion and visualisation code cells, would mark the begining of a new section in the notebook. In other cases, the assertions were typically checks to ensure that the visualisation code would not encounter any errors. This observation also holds for Arrangement 2.1, where the assertion is defined above the visualisation code, in the same cell. Common examples of such checks include checking the shape of features that are used in the visualisation. Or checking that the length of the ``x''and ``y`` features are the same to ensure a continuous line in the plot.

The signal-to-noise ratio was observed to improve in the sample of notebooks where the assertions were defined below the visualisation code. The visualisation and assertions pairs observed in Arrangment 2.2 and 3.1 were often related to each other, but were post-condition checks. Section~\ref{sec:post-cond} provides more details on post-condition checks and our rationale for excluding them from this study.

We observed the best results in notebooks where the assertion cell was not longer than 10 lines and was defined below the visualisation cell (Arrangement 3.2). We observed that most notebook authors followed the convention of writing their motivation for the visualisation in a markdown cell, before writing the code for the visualisation. The visualisations were often followed by another markdown cell to record the author's observations. And finally, the observations were translated to a analytical test using \texttt{assert} or other testing methods from external libraries. The restriction on the number of lines in the assertion cell was introduced to exclude instances where the assertion cell contained a new class definition or several helper functions that were not related to the visualisation above.

\begin{table}
  \centering
  \caption{Observations from random samples of notebooks with various
  cell arrangement.}
  \begin{tabular}{l p{0.2\textwidth} p{0.2\textwidth}}
    \hline
    \emph{\textbf{ID}}&
    \emph{\textbf{Cell arrangement}} &
    \emph{\textbf{Observations}}\\
    \hline
    1 &
    Assertion cell above visualisation cell &
    Least likely to be related to the visualisation below.\\
    \hline
    \multicolumn{3}{l}{\textbf{Assertion and visualisation in the same code cell}}\\
    \hline
    2.1 &
    Assertion above visualisation code &
    More likely to be a pre-condition check.\\
    2.2 &
    Assertion below visualisation code &
    More likely to be a post-condition check.\\
    \hline
    \multicolumn{3}{l}{\textbf{Assertion cell below visualisation cell}}\\
    \hline
    3.1 &
    Assertion cell more than 10 lines &
    Better signal-to-noise ratio compared to Arrangement 2.1. However, More likely to be post-condition check or unrelated to the visualisation above.\\
    3.2 &
    Assertion cell has max 10 lines &
    Most likely to be related to the visualisation above.\\
    \hline
  \end{tabular}
  \label{tab:cell-arrangement}
\end{table}

\subsection{Identifying Related Visualisation and Assertion Pairs}\label{sec:identify-related-pairs}

The initial sample size of 54K notebooks from Github, was reduced to the final sample size of 1.7K after applying the filters presented in Section~\ref{sec:data-collect}. We analysed all 1.7K notebooks manually to further identify 551 notebooks that contained at least one pair of visualisation and assertions that were related to each other.

We manually analysed the source code of the visualisation and assertion code cells. And excluded visualisation and assertion pairs that were not related to each other. We encountered several instances where the tracing was not straightforward due to poor coding practices. In such cases, we used one-shot prompt engineering~\cite{CITME} with the ChatGPT 4.0 model to determine if the visualisation and assertion were related to each other.

We excluded notebooks that did not train an ML model. We found several notebooks authored on topics such as linear algebra, optimisation and loss functions. Although these topics are related to ML, the visualisation and assertion pairs found in these notebooks could not be readily applied to a typically ML development workflow. Analysing such notebooks also required a working knowledge of all mathematical concepts related to ML. This was deemed beyond the scope of this study.

Finally, we excluded notebooks that were not authored entirely in English or contained faulty code that did not produce a visualisation. Table~\ref{tab:exclusion-criteria} lists all exclusion criteria used along with the number of notebooks that were excluded from this study.

\begin{table}
  \centering
  \caption{Exclusion criteria used to identify related visualisation and assertion pairs.}
  \begin{tabular}{l p{0.2\textwidth} r}
    \hline
    \textbf{ID} &
    \textbf{Exclusion Criteria} &
    \textbf{Num. excluded NBs}\\
    \hline
    \textbf{EC1} &
    The visualisation and assertion are not related &
    440\\
    \textbf{EC2} &
    Notebook does not contain a trained ML model &
    383\\
    \textbf{EC3} &
    The assertion is a post-condition check &
    200\\
    \textbf{EC4} &
    Notebook not written in English &
    159\\
    \textbf{EC5} &
    Notebook contains incorrect code &
    31\\
    \hline
    \multicolumn{2}{c}{\emph{\textbf{Total number of NBs excluded:}}} &
    1.213\\
    \hline
  \end{tabular}
  \label{tab:exclusion-criteria}
\end{table}

\subsection{Deeper Analysis of Notebooks}\label{sec:deep-dive}

We qualitatively assess the comments made for all notebooks during the analysis conducted in phase two of the study, as described in Section~\ref{sec:identify-related-pairs}.

We perform an in-depth analysis of 90 short-listed notebooks to identify unique ML testing tasks. The in-depth analysis was conducted by reading and understanding the entire contents of the notebooks in a top-down order. During the analysis, we note several data points such as the primary objective of the notebook, the type of ML problem, the type of ML model and the type of data. We manually read and comprehend each code cell in the notebooks. When present, we also read the markdown cells above and below code cells to gain further context from the notebook author. For instance, we find that markdown cells above a visualisation cell typically motivates the need for the visualisation. A markdown cell directly below a visualisation often records the observations made by the notebook author from the visualisation.

\section{Results}\label{sec:result}
\subsection{RQ1. How frequently do ML practitioners formulate analytical tests from the visualisations they create to test ML systems?}\label{sec:result-rq1}

Of the 1764 Jupyter notebooks that were analysed in this study, we found 551 notebooks that contained at least one pair of visualisation and assertions which were semantically related to each other.

Our observations indicate that AI practitioners formulate analytical tests based on the information gained from the visualisations that they create to test ML systems. However, this is an emerging testing practice since only 31\% of the notebooks analysed in this study exhibit this testing practice.

\subsection{RQ2. What are the types of visualisation and assertion pairs typically used by ML practitioners when testing ML systems?}\label{sec:result-rq2}

This study identified 26 patterns of testing using visualisations and assertions. This section highlights five interesting patterns in more detail. We plan to release the entire catalogue of patterns publicly.

\subsubsection{Decision boundary}\label{sec:svm}

\begin{figure}
  \subcaptionbox{Visualisation of the decision boundary of a SVM with a linear kernel.\label{fig:svm-linear}}{\includegraphics[width=0.49\linewidth]{../catalogue/select-04-linear.png}}
  \hfill
  \subcaptionbox{Visualisation of the decision boundary of a SVM with a RBF kernel.\label{fig:svm-rbf}}{\includegraphics[width=0.49\linewidth]{../catalogue/select-04-rbf.png}}
  \caption{Visualisation of the decision boundary of a SVM trained using different kernels.}\label{fig:svm}
\end{figure}

\begin{lstlisting}[language=Python, caption={Assertion on the accuracy of the ML model.}, label={lst:svm}]
assert accuracy_score(y_test, pred) > 0.95
\end{lstlisting}

Figures~\ref{fig:svm} presents visualisations obtained from a notebook that performs binary classification using Support Vector Machines (SVM). The model is trained on the \texttt{make\_circles}~\footnote{REFME} dataset obtained from scikit-learn~\footnote{REFME}. The visualisations show the two classes present in the training set, and the decision boundary of the model when trained first using a linear kernel (Figure~\ref{fig:svm-linear}) and later with a RBF kernel (Figure~\ref{fig:svm-rbf}).

The visualisations presented in Figure~\ref{fig:svm} contain several useful information that can be used to test and debug ML models. Humans are tuned to recognise visual patterns~\cite{CITEME}. Visualisations in Figure~\ref{fig:svm}, allows ML practitioners to quickly identify mis-classified examples. Additionally, observing the complexity of the decision boundary is a quick way to determine if the model is under or overfitting the data. For instance, in Figure~\ref{fig:svm-linear} we observe that the model is underfitting since a linear kernel too simply for the underlying data. The visualisation also allows ML practitioners to compare different ML models or variants of the same ML model to iteratively find a model that best fits the underlying training data. In this case for instance, the author of the notebook First trained the SVM with a linear kernel. And upon realising that the data is not linearly separable, chose the RBF kernel.

Listing~\ref{lst:svm} shows the corresponding assertion that was defined below the visualisations. The assertion checks if the model accuracy is above a threshold specified by the author of the notebook. The assertion is derived from the observations made from the visualisation in Figure~\ref{fig:svm-rbf}. The visulisation shows that the model is able to classify the data well which is why the notebook author sets the threshold for the accuracy in the assertion to a high value.

The assertion however, is only able to capture the information pertaining to the performance of the model. While the rest of the information that is gained from the qualitative assessment using visualisation is lost.

\subsubsection{Input Image, \texttt{predict\_proba}}

\begin{figure}
  \subcaptionbox{Random input image of a football from the test set.\label{fig:football}}{\includegraphics[width=0.49\linewidth]{../catalogue/select-35a.png}}
  \hfill
  \subcaptionbox{Random input image of tennis balls from the test set.\label{fig:tennisball}}{\includegraphics[width=0.49\linewidth]{../catalogue/select-35b.png}}
  \caption{Visualisation of random input images for a pre-trained image classifier.}\label{fig:balls}
\end{figure}

\begin{lstlisting}[language=Python, caption={Assertion below Figure~\ref{fig:football}. It checks if the probability estimate of the model is close to the specified value. The input image is of a football. Since the model is trained to detect tennis balls, the specified value is very small.}, label={lst:football}]
np.testing.assert_almost_equal(pred_proba, 2.0355723e-05, decimal=3)
\end{lstlisting}

\begin{lstlisting}[language=Python, caption={Assertion below Figure~\ref{fig:tennisball}. Compared to Listing~\ref{lst:football}, the specified value is much higher, since the input image contains tennis balls.}, label={lst:tennisball}]
np.testing.assert_almost_equal(pred_proba, 0.9988895, decimal=3)
\end{lstlisting}

Figure~\ref{fig:balls} presents visualisations obtained from a notebook that uses a pre-trained image classifier from the GluonCV library~\footnote{https://cv.gluon.ai}, to identify images that contain tennis balls. The visualisations are of two random input images, one from each class in the dataset. Listing~\ref{lst:football} presents the assertion related to Figure~\ref{fig:football}. The assertion checks if the probability estimate of the model for the given input image, is close to the specified value with a precision of up to three decimal places. Since the model is trained to detect tennis ball, and the input image contains a footbal, the specified probability estimate is very low. In contrast, Listing~\ref{lst:tennisball} specifies a much higher probability estimate since the input image contains tennis balls.

The above visualisation and assertion pair can be used to spot-check a few edge-cases or adversarial examples which the model may find difficult to classify. It is unclear however, how the specified thresholds were obtained since the authors do not provide additional explanation in the notebook. The threshold may have been derived based on the domain expertise of the ML practitioner. Another alternative could be to define them during the Requirements Engineering phase, by discussing with all involved stakeholders\cite{CITEME}.

In contrast to the visualisation and assertion pair in Section~\ref{sec:svm}, the assertion here is able to capture the author's motivation for visualising the input image along with the information obtained from it. Regardless, the notebook author still visualises the input image. This is a recurring phenomenon that we observe in other visualisation and assertion pairs in our catalogue. We discuss this phenomenon further in Section~\ref{sec:visual}.

\subsubsection{Normality}\label{sec:normal}

\begin{lstlisting}[language=Python, caption={Assertion to check that each feature in a dataset is normal. The distribution of each feature is compared to that of a normal distribution using the Kolmogorov-Smirnov test for goodness of fit from the scipy library.}, label={lst:ktest}]
for feature in range(data_transformed.shape[1]):
assert kstest(data_transformed[:, feature], 'norm').statistic < 1e-2
\end{lstlisting}

\begin{figure}
  \includegraphics[width=\linewidth]{../catalogue/select-152a.png}
  \caption{Visualisation of the distribution of a feature from the dataset.}\label{fig:kstest}
\end{figure}

The visualisation in Figure~\ref{fig:kstest} is obtained from a notebook that trains a Generative Adversarial Network (GAN) to generate data which simulates a neutrino particle detector. The notebook author transforms the training data by normalising it using the \texttt{QuantileTransform} from scikit-learn~\footnote{REFME}. After performing the transformation, the author performs a visual check by plotting the distribution of a feature from the dataset as shown in Figure~\ref{fig:kstest}. The visualisation shows a bell-shaped curve indicating that the data is normal and that the transformation worked. The author then defines the assertion presented in Listing~\ref{lst:kstest} which reflects their observation made from the visualisation.

The assertion in Listing~\ref{lst:kstest} formally verifies that all features in the dataset have a normal distribution. The assertion iterates through each column in the dataset and performs the Kolmogorov-Smirnov test to check that the distribution of the feature is similar to that of a normal distribution.

Checking the distribution of the training data is a very common task performed during the data understanding or exploratory data analysis phase of the ML development lifecycle. Several ML models assume normality in the underlying training data. The verification is often performed visually using a histogram and checking that the distribution shows a "bell-shaped" curve\cite{CITME}. However, in a production environment where ML models are continually trained using new batches of data, this assumption of normality may not always hold. The formal assertion presented in Listing~\ref{lst:kstest} makes this assumption explicit. And causes the automated ML pipeline to fail if our expectations regarding the data are no longer satisfied.

\subsubsection{Coefficient of Determination}\label{sec:r2}

\begin{lstlisting}[language=Python, caption={Assertion to check that the Coefficient of Determination ($R^2$) is higer than the specified threshold.}, label={lst:r2}]
r2_gru = r2_score(y_test, y_pred)
assert r2_gru > 0.6
\end{lstlisting}

\begin{figure}
  \includegraphics[width=\linewidth]{../catalogue/select-332a.png}
  \caption{Scatterplot of the actual and predicted labels of a dataset. The visualisation shows that there is a linear relationship between the true and predicted labels.}\label{fig:r2}
\end{figure}

Figure~\ref{fig:r2} shows a visualisation obtained from a notebook that trains a RNN to predict the levels of Nitrogen Dioxide ($NO_2$) in the next 24 hour window. The model is trained using time-series data consisting of historical readings of $NO_2$ and meteorological data such as temparate, humidity, wind speed, and such. The visualisation shows a scatterplot of the real and predicted values of $NO_2$ from the test set. The scatterplot shows a linear relationship between the ground truth and predictions indicating that the model was able to learn certain patterns in the training data. The spread of $y_{pred}$ at any given $y_{true}$ is wide. For example, at $y_{true} = 80$, the prediction can range anywhere between 30 and 120. This Indicates that the model was not able to fit the training data that well, and can be improved with fine-tuning.

Listing~\ref{lst:r2} presents the corresponding analytical test which was defined below Figure~\ref{fig:r2}. The author uses \texttt{r2\_score} metric from the scikit-learn library to quantify the performance of the model. The \texttt{r2\_score} method fits a linear regression model on the true and predicted labels and computed the Coefficient of Determination ($R^2$) based on the residuals of the linear regression model. The assertion checks that the $R^2$ score of the model is always higher than $0.6$. Since the best score possible is $1.0$, the specified threshold indicates that the author wants to ensure that $y_{pred}$ and $y_{true}$ are always linearly related to each other.

Similar to Figure~\ref{fig:svm}, the visualisation in Figure~\ref{fig:r2} allows ML practitioners to visually assess is the model is under or overfitting the training data. If the spread of $y_{pred}$ is wide, it indicates that the model is under-fitting. In contrast, if the spread is very close to the diagonal line, it indicates that the model is most likely overfitting. The visualisation can also be used to verify if there are any outliers in the predictions, for a given input. The assertion in Listing~\ref{lst:r2} however, does not capture such information. It can only check if the performance of the model is better than a specific threshold.

\subsubsection{Normalization}

\begin{figure}
  \subcaptionbox{Histogram showing the distribution of a feature before and after scaling.\label{fig:scale}}{\includegraphics[width=\linewidth]{../catalogue/select-65a.png}}
  \subcaptionbox{Histogram showing the distribution of a feature before and after normalisation.\label{fig:normal}}{\includegraphics[width=\linewidth]{../catalogue/select-65b.png}}
  \caption{Visualisation of data before and after various data pre-processing techniques.}
  \label{fig:data-pre-process}
\end{figure}

\begin{lstlisting}[language=Python, caption={Assertion to check that the mix and max of a feature fall within specified threshold derived from the visualisation presented in Figure~\ref{fig:scale}.}, label={lst:scale}]
assert scaled_data[scaled_data.argmax()] <= 1
assert scaled_data[scaled_data.argmix()] >= 0
\end{lstlisting}

\begin{lstlisting}[language=Python, caption={Similar premise as Listing~\ref{lst:scale}, however this assertion is based on Figure~\ref{fig:normal}.}, label={lst:normal}]
assert normalized_data[normalized_data.argmax()] < 10
assert normalized_data[normalized_data.argmin()] > -7
\end{lstlisting}

Visualisations presented in Figure~\ref{fig:data-pre-process} were obtained from a notebook that presents a tutorial on scaling and normalization---two common data pre-processing techniques in ML. The visualisations show a comparison of the data before and after each pre-processing technique is applied to a feature from the data. The author scales the data using the \texttt{MinMaxScaler} from the scikit-learn library~\footnote{REFME}, between $[0, 1]$. The author performs a visual check, by comparing the distribution of the data before and after scaling, using a histogram (as shown in Figure~\ref{fig:scale}). Similarly, the author normalises the data using the \texttt{PowerTransformer} class from scikit-learn~footnote{REFME}, and performs a visual check as shown in Figure~\ref{fig:normal}.

Listing~\ref{lst:scale}~and~\ref{lst:normal} show the assertions that were defined below Figure~\ref{fig:scale}~and~\ref{fig:normal} respectively. Both assertions check that the min and max of the feature is within the expected bounds after applying the respective pre-processing technique. The assertion in Listing~\ref{lst:scale} is felt to be unneccessary since the min and max thresholds are set to the theoretical limits. The assertions in Listing~\ref{normal} are more relevant, however it is not clear why the author include such a wide margin in the thresholds.

\section{Implications}\label{sec:discuss}

This study mined 54K Jupyter Notebooks that contained an assertion. Further analysis identified 1.7K notebooks which contained an assertion directly below a visualisation. Of the 1.7K notebooks, 551 notebooks contained a visualisation and assertion pair which were semantically related to each other.A deeper analysis of 90 Jupyter notebooks revealved 26 unique ML testing tasks which utilised a combination of visualisations and assertions.

The results from RQ2 provide emprical evidence to support the use of visualisations to test various properties of ML systems.  Visualisations are an efficient and effective medium to present complex information, that is easier to interpret by humans with the aid of visual cues. Visualisations used to test ML system properties however, demand manual verification effort. Information gained from visualisations and subsequent decisions made based on this information, must be manually verified in every iteration of the ML development cycle.

% NOTE link to accumulation of technical debt due to highly experimental nature, and constant change in the system

Assertions or analytical tests extracted from ML visualisations can reduce this manual verification effort. Assertions encapsulate the knowledge gained from the visualisation with regards to the property of the ML system which is currently under test. Formal assertions document the observations made by the AI practitioner regarding the model or the data, at a specific point in time. These observations should not be violated by the future model development cycles unless the requirements of the ML system are changed. Assertions scale across organizational or personnel changes as future AI practitioners can refer to the assertions to understand how the visualisations were previously interpreted.

Results for RQ1 indicate that combining ML visualisations with analytical assertions is an emerging testing practice within the ML community. However in results from RQ2, we see several examples of assertions that fail to capture all the information gained from its visual counter-part. Our observations indicate a need to bridge the gap between qualitative assessment of ML systems using visualisations and deriving analytical assertions from the visualisations. We find the need for automated tools that make this translation process easier for AI practitioners.

\section{Threats to Validity}\label{sec:threats}
% defend use of github as source for notebooks:
% - why did kaggle not work (refer to notes)
% - why did the existing replication packages not work?

% TODO defend the use of a simple search query on github
Our approach also prevents the need to craft a custom search query based on an exhaustive list of keywords that appear in all python testing libraries. We intentionally keep the search criteria on Github broad by only checking for the appropriate filetype and the presence of the ``assert'' keyword. This is because, the ``assert'' keyword is used in python to define tests. Additionally, popular python libraries which provide a testing interface (such as nose, unittest and numpy), contain the same keyword in their API.

% TODO don't mention quantity of random sample; then we need to explain/defend this choice
We used an iterative approach to determine the effectiveness of each filter that was used. We applied each filter in the same order as presented in Table~\ref{tab:filters}, starting with F1. We randomly sampled 50 notebooks from the population of notebooks obtained after applying the filter. We then conduct the qualitative assessment presented in Section~\ref{REFME} and observe the number of related visualisation and assertion pairs. In each new iteration, we progressively add the other filters and include the filter if the number of relevant visualisation and assertion pairs increases.

% TODO defend use of one-shot prompt engineering for identifying related visualisation and assertion code cells; how did we mitigate this threat?

\section{Conclusion}\label{sec:conclude}

\bibliographystyle{IEEEtran}
\bibliography{bibliography}

\end{document}
\endinput