### 2024-01-31 report/bsc-research-project-proposal.qmd
refactor: migrate shome2023llm files
It does not make sense to keep shome2023llm as a separate project
since its a sibling of shome2023notebook.

This commit moves all relevant files of shome2023llm into this repo.

---
title: "Bsc. Research Project Proposal"
author: "Arumoy Shome"
format:
  pdf:
    mainfont: "Iosevka Aile"
    monofont: "Iosevka"
bibliography: bibliography.bib
csl: acm-sig.csl
---

# Introduction

Visualisations are employed at various stages of a Machine Learning
(ML) pipeline. They are used to understand and verify data properties
during the early stages, summarise metrics and fine-tune models during
development, and monitor performance post-deployment. The iterative
and experimental nature of building ML systems heavily relies on
insights from visualisations to guide design and implementation
decisions
[@yuan2020survey;@hohman2019visual;@amershi2015modeltracker;@wexler2019what-if;@shome2022data;@haakman2021ai].

However, real-world data that ML systems encounter post-deployment
seldom remain static. They often change as a reflection of the world,
potentially violating initial assumptions made during development
[@amershi2019software;@lwakatare2021on]. Every subsequent iteration of
the ML development cycle used to retrain and update the ML model,
therefore demands manual validation of the visualisations that were
used to test ML system properties.

Assertions or analytical tests derived from ML visualisations can
significantly reduce manual verification efforts. Such formal
assertions record the AI practitioner's observations about the model
or data at a specific moment. They also serve as a reference point for
future AI practitioners to understand the interpretations made from
earlier visualisations.

In a prior study, we mined $54,070$ Jupyter notebooks from Github and
created a high-quality dataset of $269$ semantically related
visualisation-assertion (VA) pairs. The input feature space comprises
of a rich source of information comprising of visualisations, Python
source code, and associated markdown text.

This research project will focus on developing automated tools using
state-of-the-art Large Language Models (LLMs) that aid ML
practitioners when working with visualisations. The research questions
for this project are categorised into 3 distinct tracks and presented
below.

# Track 1: Exploring Multi-Model Input Feature Space for Translating ML Visualisation to Python Assertions

This track will focus on exploring various multi-model input feature
space for the LLMs. Students will work with various open-source LLMs
from the HuggingFace library and develop an optimal training set that
produces the best results. There are 3 research questions here:

+ **RQ1**: Can image-to-text models translate visual properties
  observed in ML visualisation to Python assertions?

  The objective here is to experiment with various image-to-text
  models with a dataset comprising of PNG images containing
  visualisations obtained from Jupyter Notebooks.

+ **RQ2**: Does a multi-modal dataset improve the results of the AI
  models?

  Here we will augment the dataset with additional features extracted
  from the Python code that produces the visualisation and the
  surrounding Markdown text written by the notebook authors.

+ **RQ3**: Do LLMs4Code models perform better compared to
  general-purpose LLMs?

  There are several LLMs that have been trained on a large corpus of
  code written in programming languages. The objective of RQ3 is to
  compare such models with general purpose LLMs and determine if they
  are better suited for the task.

# Track 2: Detecting "Smelly" Python Visualisation Code

This track will focus on creating an automated tool that can detect
code smells in the Python code that generate ML visualisations.

+ **RQ4**: What are the most frequently occuring code smells in ML
  visualisation code written in Python?

  The focus here will be to propose code smells observed in Python
  code that generate visualisations. This track is more open-ended as
  students can propose code smells from different aspects such as
  design of the visualisation and performance of the underlying code.

# Track 3: Automatically Detecting "Related" VA Pairs in Jupyter Notebooks

Creating a dataset of semantically related VA pairs requires a lot of
manual effort. This track will focus on exploring strategies to
automate this process.

+ **RQ5**: Can code similarity metrics automatically detect related
  VA pairs?

  In this RQ the focus will be on experimenting with various code
  similarity metrics [@eghbali2022crystalbleu] which can automatically
  detect if a given VA pair are related to each other.

# References

::: {#refs}
:::