### 2024-07-29 docs/index.qmd
feat: init online appendix

---
title: Understanding Feedback Mechanisms in Machine Learning Jupyter Notebooks
subtitle: Online Appendix
abstract: |
  This website hosts the online appendix for the research paper titled
  *Understanding Feedback Mechanisms in Machine Learning Jupyter
  Notebooks* which has been submitted to the Journal of Empirical
  Software Engineering.
toc: true
format:
  html:
    theme: cosmo
website:
  site-url: "https://arumoy.me"
  back-to-top-navigation: true
page-footer:
  center: "All materials on this website are open sourced under the CC-BY 4.0 license."
---

# (RQ2) How is explicit feedback from assert statements used to validate ML code written in Jupyter notebooks?
## Data Shape Check ($N = 26$)

| Key | Code                                                                                                        |
|-----|-------------------------------------------------------------------------------------------------------------|
| A5  | `assert y_valid.shape == (1132,)`                                                                           |
| A17 | `assert X.shape[1] == 13, 'Did you drop/lose some columns in X? Did you properly load and split the data?'` |
| A29 | `assert len(test_y_preds) == len(test_y), 'Unexpected number of predictions.'`                              |
| A31 | `assert img.shape == (112, 92)`                                                                             |
| A76 | `assert len(encoding['token_type_ids']) == max_seq_length`                                                  |
| A84 | `assert red.get_shape().as_list()[1:] == [224, 224, 1]`                                                     |
| A90 | `assert len(X_train) == 2000`                                                                               |
| A93 | `assert temp_embed.shape[0] == stride`                                                                      |

## Data Validation Check ($N = 14$)

| Key | Code                                                                                |
|-----|-------------------------------------------------------------------------------------|
| A41 | `assert np.all(np.unique(X['smoke'].values) == np.array([0, 1]))`                   |
| A44 | `assert np.all(np.unique(X['smoke'].values) == np.array([0, 1]))`                   |
| A46 | `assert np.isclose(stdev_norm, 1.0, atol=1e-16)`                                    |
| A52 | `assert grouped_users['user_id'].nunique() == user_engagement['user_id'].nunique()` |
| A65 | `assert np.all(y <= nb_classes)`                                                    |
| A73 | `assert df['clf'].value_counts()[1] == len(df[df['quality'] >= 7])`                 |

## Model Performance Check ($N = 11$)

| Key | Code                                                                                                                     |
|-----|--------------------------------------------------------------------------------------------------------------------------|
| A7  | `assert len(neighbours_1) == 20, "Neighbors don't match!"`                                                               |
| A15 | `assert np.allclose(verify('images/camera_1.jpg', 'bertrand', database, FRmodel), (0.54364836, True))$`                  |
| A19 | `assert np.allclose(linear_model.coef_, [[1.57104472, 0.92521608]]), 'The model parameters you learned seem incorrect!'` |
| A38 | `assert 0.75 < auc(fpr, tpr) < 0.85`                                                                                     |
| A58 | `assert np.isclose(accuracy, 0.9666666666666667)`                                                                        |

## Existence Check ($N = 8$)

| Key | Code                                                                                 |
|-----|--------------------------------------------------------------------------------------|
| A23 | `assert np.all(orders.groupby('user_id') .days_since_prior_order.tail(1).notnull())` |
| A42 | `assert not lab_s.isnull().values.any()`                                             |
| A43 | `assert len(data) != 0, 'cannot divide by zero'`                                     |
| A50 | `assert not np.any(np.isnan(X))`                                                     |
| A51 | `assert data.target.notnull().all()`                                                 |
| A63 | `assert X.isnull().sum().sum() == 0`                                                 |
| A79 | `assert not processed_data_df.isna().any().any()`                                    |
| A86 | `assert p0 in poi_info.index`                                                        |

## Resource Check ($N = 7$)

| Key | Code                                                                                                                                            |
|-----|-------------------------------------------------------------------------------------------------------------------------------------------------|
| A10 | `assert le_path.is_file(), f"Label encoder file not found at {le_path}. Make sure 'label_encoder.pkl' exists in the lightning_logs directory."` |
| A14 | `assert self.model is not None, 'Model is not loaded, load it by calling .load_model()'`                                                        |
| A18 | `assert pd.__version__.rpartition('.')[0] == '1.0', f"Unexpected pandas version: expected 1.0, got {pd.__version__.rpartition('.')[0]}"`        |
| A37 | `assert svm.fit_status_ == 0, 'Forgot to train the SVM!'`                                                                                       |
| A60 | `assert f2.gca().has_data()`                                                                                                                    |
| A67 | `assert pm.__version__ == '3.9.2'`                                                                                                              |
| A74 | `assert os.path.exists(image_dir)`                                                                                                              |

## Type Check ($N = 5$)

| Key | Code                                                                                          |
|-----|-----------------------------------------------------------------------------------------------|
| A2  | `assert isinstance(X_trn, torch.FloatTensor), 'Features should be float32!'`                  |
| A35 | `assert isinstance(column_transformer, ColumnTransformer), "Input isn't a ColumnTransformer"` |
| A40 | `assert isinstance(model_3, sklearn.ensemble.RandomForestClassifier)`                         |
| A81 | `assert is_all_ints(filled_df[r]) is True`                                                    |
| A88 | `assert isinstance(betas, np.ndarray)`                                                        |

## Mathematical Property Check ($N = 4$)

| Key | Code                                                                     |
|-----|--------------------------------------------------------------------------|
| A3  | `assert (xH - wH) % self.stride == 0`                                    |
| A25 | `assert test_output.std() < 0.15, "Don't use batchnorm here"`            |
| A56 | `assert np.allclose(e_v_states[:, -1], np.ones_like(e_v_states[:, -1]))` |
| A64 | `assert np.allclose(T, T.T)`                                             |

## Batch Size Check ($N = 3$)

| Key | Code                                                                                                                                        |
|-----|---------------------------------------------------------------------------------------------------------------------------------------------|
| A21 | `assert x.size(0) % batch_size == 0, f'the first dimension of input tensor ({x.size(0)}) should be divisible by batch_size ({batch_size})'` |
| A28 | `assert image_size % patch_size_small == 0, 'Image dimensions must be divisible by the patch size.'`                                        |
| A70 | `assert n_img > batch_size`                                                                                                                 |

## Network Architecture Check ($N = 3$)

| Key | Code                                                                                   |
|-----|----------------------------------------------------------------------------------------|
| A11 | `assert self.encoder_conv_01[0].weight.size() == self.vgg16.features[2].weight.size()` |
| A62 | `assert self.encoder_conv_01[0].weight.size() == self.vgg16.features[2].weight.size()` |
| A75 | `assert reg in ['none', 'l2']`                                                         |

## Data Leakage Check ($N = 1$)

| Key | Code                                                                                |
|-----|-------------------------------------------------------------------------------------|
| A33 | `assert len(set( tr_df.PetID.unique()).intersection(valid_df.PetID.unique())) == 0` |

# (RQ3) How is implicit feedback from print statements and last cell statements used when writing ML code in Jupyter notebooks?

## Model Performance Check ($N = 33$)

| Key | Code                                                                                                 |
|-----|------------------------------------------------------------------------------------------------------|
| P3  | `print('The mean accuracy with 10 fold cross validation is: %s ' % round(scores * 100, 2), '%')`     |
| P6  | `print('RMSE:', np.sqrt(metrics.mean_squared_error(y_test, pred)))`                                  |
| P18 | `print('The Accuracy is:', accuracy_score(y_test, y_pred))`                                          |
| P50 | `print('Classification Report: SVM (validation data)')$`                                             |
| P54 | `print('Intercept value:', lm.intercept_)`                                                           |
| L3  | `skplt.metrics. plot_confusion_matrix(Y_val, Vote.predict(X_val), normalize=True, figsize=(10, 10))` |
| L52 | `spot_check_recs(classifier, 910)`                                                                   |

## Data Distribution ($N = 7$)

| Key | Code                                                                                |
|-----|-------------------------------------------------------------------------------------|
| L2  | `_ = sns.catplot(x='category_id', y='likes', data=train, height=5, aspect=1.5)`     |
| L9  | `sns.kdeplot(data=data.loc[ data['Survived'] == 0].Age, label='Died', shade=True)`  |
| L14 | `pd.pivot_table(train, index='Survived', values=['Age', 'SibSp', 'Parch', 'Fare'])` |
| L25 | `sns.countplot(house_pred['OverallQual'])`                                          |
| L48 | `x_train.describe()`                                                                |

## Resource Check ($N = 7$)

| Key  | Code                                      |
|------|-------------------------------------------|
| P68  | `print('GPU is available')`               |
| P71  | `print('Hub version: ', hub.__version__)` |
| P82  | `print('Running on TPU ', tpu.master())`  |
| P86  | `print('Cuda is available')`              |
| P107 | `print('Model loaded')`                   |
| L64  | `full_table.head(-5)`                     |
| L66  | `prostate_cancer_df.shape`                |

## Spot Check ($N = 5$)

| Key  | Code                          |
|------|-------------------------------|
| L60  | `X_pca.head()`                |
| P64  | `print(np.max(cur[:, :, 1]))` |
| P114 | `print(onehot_encoded)`       |

## Model Training Check ($N = 4$)

| Key | Code                                                                                                                                            |
|-----|-------------------------------------------------------------------------------------------------------------------------------------------------|
| L8  | `autoencoder.fit(x=X_train, y=X_train, epochs=15, validation_data=[X_test, X_test], callbacks=[keras_utils.TqdmProgressCallback()], verbose=0)` |
| L31 | `adaBoost.fit(X_train, y_train)`                                                                                                                |
| L42 | `m_r.best_params_`                                                                                                                              |

## Missing Value Check ($N = 3$)

| Key | Code                                                                           |
|-----|--------------------------------------------------------------------------------|
| P74 | `print(train_df.isnull().sum())`                                               |
| L12 | `sns.heatmap(test_df.isnull(), yticklabels=False, cbar=False, cmap='viridis')` |
| L36 | `test.isna().sum().unique()`                                                   |

## Shape Check ($N = 3$)

| Key  | Code                                                      |
|------|-----------------------------------------------------------|
| P4   | `print('no.of examples in test data : ', len(test_data))` |
| P32  | `print('Training set shape : ', x_train.shape)`           |
| P117 | `print('Y_train.shape: ', Y_train.shape)`                 |

## Data Relationship Check ($N = 2$)

| Key | Code                                                                                                      |
|-----|-----------------------------------------------------------------------------------------------------------|
| L6  | `b = sns.relplot(x='SIZE', y='Cash', hue='CLARITY', alpha=0.9, palette='muted', height=8, data=raw_data)` |
| L10 | `sns.regplot(x='X4 number of convenience stores', y='Y house price of unit area', data=data)`             |

## Type Check ($N = 2$)

| Key | Code                                |
|-----|-------------------------------------|
| P43 | `print('data type:', images.dtype)` |
| L71 | `type(Y)`                           |

## Execution Time Check ($N = 1$)

| Key | Code                       |
|-----|----------------------------|
| P66 | `print('Total Run Time:')` |

## Network Architecture Check ($N = 1$)

| Key | Code               |
|-----|--------------------|
| P92 | `print(MyNetwork)` |